A re-examination of the role of friction in the original Social Force Model

### 2019-06-14 paper.tex
retoques esteticos
hice casi todas las correcciones que me pidio guille. Falta revisar el parrafo de las citas de los ultimos 10 años

\textcolor{red}{In the last decades, many microscopic models 
for crowd dynamics have been developed. The first model was developed by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models represent the interaction between 
pedestrians and their environment using different kinds of forces inspired in physical systems. For 
example, Okazaki (1979) modeled the movement the pedestrians as the motion 
of a magnetized object within a magnetic field \cite{Okazaki}.}\\

@@

\textcolor{red}{More recent measures of the fundamental diagram in an extreme density
 situation show that, in some situations, the flux may increase with the density (Ref.~\cite{lohner1}). 
This seems to contradict the results of Ref.~\cite{helbing3}. Nevertheless, the measurements of Ref.~\cite{lohner1}
were taken in the Kabba while the measurements of Ref.~\cite{helbing3} were taken at the entrance of
the Jamaraat bridge. As stated in Ref~\cite{lohner1}, the fundamental diagram may depend
strongly on the type of motion and on the nature of the event. Pedestrians in the Kaaba
may walk differently than pedestrians entering to the Jamarat bridge 
(due to differences according to the moment of the ritual).}\\

\textcolor{red}{A lot of research has been done using data from the Mecca pilgrimage.
Although in this work we focused on Ref.~\cite{helbing3}. There were several investigations 
using real data to calibrate parameters of the PedFlow model developed by 
Lohner et al. Ref.~\cite{lohner2}. Recently a paper was published that uses 
the Kalman filter ensemble to adjust the parameters of this model Ref.~\cite{Togashi1}. 
For more information on PedFlow and Mecca pilgrimage measurements
see Refs.~\cite{Boltes,Dridi1,Dridi2,Baqui1,lohner1} }\\

@@

We show that the experimental data (say, the flux-density diagram) can be 
modeled under quasi-stationary conditions in the high density regime. Our 
starting point is the re-examination of the (basic) SFM.
\textcolor{red}{We proposed to reduce in units the equation of motion of 
the SFM (Eq.~\ref{eqn_appendix_1} and Eq.~\ref{eqn_appendix_2}). This leads us
to two control parameters} $\mathcal{A}$ and $\mathcal{K}$ (see Section~\ref{appendix_1}) that may handle
the collective motion in dense crowds. We presume that physical contact is
a key feature in dense crowds, despite the fact that other issues may
also contribute to the flow reduction \cite{johansson1}. However, the latter could
be satisfactorily omitted in past research \cite{johansson}, and thus, we will not
attempt to introduce further extensions to the (basic) SFM for the
sake of simplicity.\\

@@

We first noticed that low-density situations ($\rho<5$) lead to a cruising velocity profile $v=v_d$. This is valid for every location in the corridor (not only the center as was previously noticed in Section \ref{fundamental_diagram}). For higher densities($\rho>5$), the velocity profile turns into a parabola-like function. This shape resembles the usual velocity profile for laminar flow in a viscous fluid, where the velocity increases toward the center of a tube. In our case, pedestrians near the walls are the ones with the lower velocity. The velocity increases when departing from the wall until it reaches the maximum at the center of the corridor. This behavior suggests that the wall friction on the pedestrians, is playing a relevant role on the velocity distribution. \textcolor{red}{The velocity profile is in accordance with the empirical measurements of Ref.~\cite{zhang1}.}  \\

@@

\textcolor{red}{We interpret $\mathcal{A}$ as a parameter related exclusively to the 
repulsion between pedestrians and $\mathcal{K}$  as parameter related to the friction-repulsion.
While the former is valid in every situation, the latter only takes part when the
pedestrians are in contact.}\\

@@

We found two unexpected results. Increasing the density produces bigger size clusters until the size distribution suddenly switches to a bimodal distribution (compare \ref{size_distribution_w22_density5} with \ref{size_distribution_w22_density5_5} and \ref{size_distribution_w22_density4.5_kx10} with \ref{size_distribution_w22_density5_kx10}). Once
this phenomenon occurs, any of two possibilities may
appear: the pedestrian belongs to a small cluster (out
of many caged in the crowd) or he (she) belongs to the
giant cluster (with a size comparable to the entire crowd).
\textcolor{red}{In other words, the bimodal distribution occurs because when 
the giant component is formed there are some small clusters ``caged"
inside the giant component. These small clusters do not touch any 
pedestrian of the giant component.} \\

@@

>>>>

\textcolor{red}{In the last decades, many microscopic models 
for crowd dynamics were developed. The class of force-based model was introduced by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models represent the interaction between 
pedestrians and their environment using different kinds of forces inspired in physical systems. For 
example, Okazaki (1979) modeled the movement of pedestrians as 
a magnetized object immersed a magnetic in field \cite{Okazaki}.}\\

@@

\textcolor{red}{More recent measures of the fundamental diagram in an extremely dense
 situation show that, sometimes, the flux may increase with the density (Ref.~\cite{lohner1}). 
This seems to contradict the results of Ref.~\cite{helbing3}. But both results seem 
not to be completely  comparable since data was acquired at different points, and thus,
at different stages of the ritual. Ref.~\cite{lohner1}, indeed, stated that the fundamental 
diagram may strongly depend on the type of motion and the nature of the event.}\\

\textcolor{red}{A lot of research has been done using data from the Mecca pilgrimage.
Although in this work we focused on Ref.~\cite{helbing3}. There were several investigations 
using real data to calibrate parameters of the PedFlow model developed by 
L\"ohner et al. Ref.~\cite{lohner2}. Recently a paper was published that uses 
the Kalman filter ensemble to adjust the parameters of this model Ref.~\cite{Togashi1}. 
For more information on PedFlow and Mecca pilgrimage measurements
see Refs.~\cite{Boltes,Dridi1,Dridi2,Baqui1,lohner1} }\\

@@

We show that the experimental data (say, the flux-density diagram) can be 
modeled under quasi-stationary conditions in the high density regime.\\

\textcolor{red}{Our starting point is the re-examination of the (basic) SFM. 
We propose a reduce-in-units equation of motion of the SFM. The social 
force $\mathbf{f}_s$ depends on an intensity $\mathcal{A}$ while the granular force $\mathbf{f}_g$ depends
on $\mathcal{K}$. In Section~\ref{appendix_1} we derivate this reduce-in-units equation of motion
and its corresponding control parameters. }\\

%Our starting point is the re-examination of the (basic) SFM.
%\textcolor{red}{We propose a reduce-in-units equation of motion of 
%the SFM (Eq.~\ref{eqn_appendix_1} and Eq.~\ref{eqn_appendix_2}). This leads us
%to two control parameters} $\mathcal{A}$ and $\mathcal{K}$ (see Section~\ref{appendix_1}) that may handle
%the collective motion in dense crowds.

 We presume that physical contact is
a key feature in dense crowds, despite the fact that other issues may
also contribute to the flow reduction \cite{johansson1}. However, the latter could
be satisfactorily omitted in past research \cite{johansson}, and thus, we will not
attempt to introduce further extensions to the (basic) SFM for the
sake of simplicity.\\

@@

We first noticed that low-density situations ($\rho<5$) lead to a cruising velocity profile $v=v_d$. This is valid for every location in the corridor (not only the center as was previously noticed in Section \ref{fundamental_diagram}). For higher densities($\rho>5$), the velocity profile turns into a parabola-like function. This shape resembles the usual velocity profile for laminar flow in a viscous fluid, where the velocity increases toward the center of a tube. In our case, pedestrians near the walls are the ones with the lower velocity. The velocity increases when departing from the wall until it reaches the maximum at the center of the corridor. This behavior suggests that the wall friction on the pedestrians, is playing a relevant role on the velocity distribution. \textcolor{red}{The velocity profile is in agreement with the empirical data in Ref.~\cite{zhang1}.}  \\

@@

\textcolor{red}{We envisage $\mathcal{A}$ as a parameter related exclusively to the 
repulsion between pedestrians and $\mathcal{K}$  as parameter related to the friction-repulsion.
While the former is valid in every situation, the latter only takes part when the
pedestrians are in contact.}\\

@@

We found two unexpected results. Increasing the density produces bigger size clusters until the size distribution suddenly switches to a bimodal distribution (compare \ref{size_distribution_w22_density5} with \ref{size_distribution_w22_density5_5} and \ref{size_distribution_w22_density4.5_kx10} with \ref{size_distribution_w22_density5_kx10}). Once
this phenomenon occurs, any of two possibilities may
appear: the pedestrian belongs to a small cluster (out
of many ``caged" in the crowd) or he (she) belongs to the
giant cluster (with a size comparable to the entire crowd).
\textcolor{red}{In other words, the bimodal distribution occurs because,
after the giant component is formed, many small clusters remain ``caged" 
inside the giant component. These small clusters (or single individuals)
do not touch permanently any pedestrian belonging to the giant component.} \\

@@

%\appendix
